---
title: "Banknotes in Circulation"
author: "Victor Cuspinera"
date: "23/10/2020"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

One of the main activities of Banco de Mexico is to issue new Banknotes and Coins and ensure an adequate level of Banknotes and Coins in Circulation to satisfy the Cash requirements of the people, businesses and companies, to allow the Economy of the country work properly.

The aim of this document is to analyze the Banknotes in Circulation using the SIE API with the `siebanxicor` R-package, and the custome function `sie_function` from the _SIE_function.R_ script.

## Quick review

I use the developed customed function for an overview of the first draft to look into the Banknotes in Circulation by denomination.

```{r quick review}
# load libraries
library("siebanxicor")
library("tidyverse")
library(reshape2)

# call and set the token
token_file <- read.csv("../token/SIE_Token.csv", header=FALSE)
setToken(token_file$V2)

# call the customed function from an RScript
source("SIE_function.R")

# setting the variables for current Notes in Circulation
my_series <- c('SM1472', 'SM1478', 'SM1479', 'SM1480', 'SM1481', 'SM1482')
my_names <- c('20','50','100','200','500','1000')
my_title <- "Notes in Circulation"
my_start <- '2015-01-01'

my_route <- "../img/"

# run the function
series_NIC <- sie_function(my_series, my_names, 
                           my_title, route = my_route,
                           y_lab = "Million of banknotes",
                           startDate = my_start)
```

And this is an example of how does the Notes in Circulation (NIC) tidy data frame looks like:
```{r tidy df, echo = FALSE}
# print an example of the data frame
series_NIC %>% tail()
```

## Getting a squared data frame
Tidy data is not always the easiest way to look at through the human-eye. The next code chunk wranges the previously obtained tidy data frame into a messy and square data frame with one column per denomination.
```{r messy df}
# get square data frame
squared_NIC <- dcast(series_NIC, date~serie)

# rename the series
names(squared_NIC) <- c("date", my_names)

# order by date
squared_NIC <- squared_NIC[order(as.Date(squared_NIC$date, format = "%Y-%m-%d")),]

# print an example of the data frame
squared_NIC %>% head()
```

## Looking into the annual growth
It would be also useful to look into the annual growth (comparing the number of NIC of a specific month vs. the data of the same month from the previous year).
```{r annual growth}
# get the growth rate
growth_NIC <- series_NIC %>%
    group_by(serie) %>%
    mutate(annual_growth = round((value - lag(value, 12)) * 100 / lag(value, 12), 1)) %>% 
    subset(select = -c(value)) %>%
    drop_na()

# print an example of the data frame
growth_NIC %>% tail()

# plot
growth_NIC %>% ggplot() +
  geom_line(aes(x = date, y = annual_growth, color = serie_name), size = 0.7) +
  labs(title = "Growth of Banknotes in Circulation", subtitle = "2016-2020", x = "", y = "Percentage (%)", color = "Denomination") +
  scale_color_manual(values = c("dodgerblue3", "pink2", "firebrick1", "chartreuse3", "gold", "darkorchid3")) +
    annotate(geom = "rect", xmin = as.Date("2020-03-01", "%Y-%m-%d"), xmax = as.Date(Sys.Date(), "%Y-%m-%d"), ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.15) +
  annotate("text", x = as.Date("2020-02-01", "%Y-%m-%d"), y = 29.5, label = "Covid-19 \nlockdown", size = 4, color = "darkred", alpha = 0.8) + 
  theme_bw()

# save plot
ggsave(paste0(my_route, "Growth of Banknotes in Circulation.png"))
```

## Cumulative growth over the year
While the previous graph shows the annual growth of Banknotes in Circulation, a level graph of the cumulative growth of Banknotes in Circulation show us the need of cash during the year. 

The first step would be get the Banknotes in Circulation per year.
```{r total NIC per year}
# get the growth rate
df_total_NIC <- series_NIC %>%
  group_by(date) %>%
  summarise(NIC = sum(value)) %>%
  mutate(year = format(as.Date(date, format = "%Y-%m-%d"),"%Y")) %>%
  mutate(month = format(as.Date(date, format = "%Y-%m-%d"),"%b"))

# set order of months
df_total_NIC$month = factor(df_total_NIC$month, levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))

# show data.frame
df_total_NIC %>% head()

# plot
df_total_NIC %>% ggplot(aes(x = month, y = NIC, color = year, group = year))+
  geom_line() +
  geom_point() + 
  labs(title = "Banknotes in Circulation", subtitle = "2015-2020, per year", x = "Month", y = "Millions of pieces", color = "Year") +
  theme_bw()

# save plot
ggsave(paste0(my_route, "Total Banknotes in Circulation per year.png"))
```

The nexyt step is obtain the growth rate using January as starting point.
```{r NIC cumulative growth, warning=FALSE}
# fix the initial value for each year
df_initial <- df_total_NIC %>%
  #mutate(NIC_diff = NIC - lag(NIC) ) %>%
  melt(id = c("month", "year"), measure="NIC") %>%
  dcast(month ~ year) %>%
  subset(select = c("month", "2016", "2017", "2018", "2019", "2020")) %>% slice(1)
df_initial <- as.numeric(df_initial[1,2:6])

# get the difference for the value reported in each month
#    compared with the January value of the same year
df_diff_NIC <- df_total_NIC %>%
  mutate(NIC_diff = NIC - lag(NIC) ) %>%
  melt(id = c("month", "year"), measure="NIC_diff") %>%
  dcast(month ~ year) %>%
  subset(select = c("month", "2016", "2017", "2018", "2019", "2020"))
the_months <- df_diff_NIC[[1]]
the_columns <- colnames(df_diff_NIC)
df_diff_NIC[1,2:6] <- 0
df_diff_NIC <- df_diff_NIC[,2:6] %>% 
  apply(2, cumsum) %>% matrix(ncol=5) 


# get the growth dataframe
my_matrix <- sweep(df_diff_NIC, 2, df_initial, "/") * 100
df_growth_NIC <- data.frame(the_months, my_matrix) %>%
  rename_all(~the_columns) %>%
  melt(id=c("month"))

# plot
df_growth_NIC %>% ggplot(aes(x = month, y = value, color = variable, group = variable)) +
  geom_line() +
  geom_point() +
  labs(title = "Cumulative Growth of Banknotes in Circulation",
       subtitle = "2016-2020",
       x="Month", y="Percentage (%)", 
       color = "Year") +
  theme_bw()

# save plot
ggsave(paste0(my_route, "Cumulative Growth of Banknotes in Circulation.png"))
```

