---
title: "Diverse examples of the custom `sie_function`"
author: "Victor Cuspinera"
date: "25/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The aim of this document is to share diverse examples on how to use the custome function `sie_function`, developed to get a quick view of diverse selected SIE series.

For details of the code of this function, look at the Rscript [SIE_function.R](https://github.com/vcuspinera/SIE_API_guide/blob/main/src/SIE_function.R) in the **src** folder of this repository.

## SIE's Series Catalogue

This notebook shows diverse examples using different time series published by Banco de Mexico in the SIE, using series as exhange rates, operations and number of ATMs, retail transactions, among other series.

[Click here](https://www.banxico.org.mx/SieAPIRest/service/v1/doc/catalogoSeries) to look for the complete catalogue of the SIE's time series published by Banco de Mexico.

## Examples

The time series used in this notebook for examples are:

1. Exchange Rate, of U.S. Dollar, Canadian Dollar, British Pound, Australian Dollar, and Euro to Mexican Pesos, since 2000.
2. Operations in ATMs with Debit and Credit Cards.
3. Number of ATM for the eight most-populated States in Mexico, since 2010.
4. Retail payment systems transactions at ATMs, POS, Checks, and Transfers by Electronic Payments, Internet Banking and Phone, from 2016 to 2019.
5. Consumer Price Index (INPC), main index and subindexes, during Enrique Peña Nieto's Presidency of Mexico (2012-2018).

### 0. Load libraries and set token
The first step is to use the custome function is load the library, as well to call and set the token.

```{r load library and set token, warning = FALSE}
# load libraries
library("siebanxicor")
library("tidyverse")

# bring the token
token_file <- read.csv("../token/SIE_Token.csv", header=FALSE)

# set the token
setToken(token_file$V2)

# call the customed function from the RScript
source("SIE_function.R")
```


### 1. Exchange Rate
Exchange rate of U.S. Dollar, Canadian Dollar, British Pound, Australian Dollar, and Euro to Mexican Pesos, since 2000.

```{r exchange rate}
serie_tc <- c("SF60653", "SF57771", "SF57815", "SF57753", 
              "SF57923")
name_tc <- c("USD", "CAD", "GBP", "AUD", "EUR")
title_tc <- "Exchange Rates"
my_start <- '2000-01-01'
my_y <- "Mexican Pesos"
my_x <- "Date"

# run the function
df_tc <- sie_function(serie_tc, name_tc, 
            title_tc, route="../img/",
            y_lab = my_y, x_lab = my_x,
            startDate=my_start)
```

### 2. Operations in ATMs
Series with the number of operations in ATMs with Debit and Credit Cards.
```{r atm operations}
serie_atm <- c("SF62270", "SF62271")
name_atm <- c("with Debit Card", "with Credit Card")
title_atm <- "Operations in ATMs"
my_y <- "Number of operations"
my_x <- "Quarter"

# run the function
df_atm <- sie_function(serie_atm, name_atm, 
            title_atm, route="../img/",
            y_lab = my_y, x_lab = my_x)
```


### 3. Number of ATM for the most populated States 
Quarterly time series of the number of ATMs by State, for the 8 states more populated in Mexico, since 2010.

```{r atm number}
serie_state <- c("SF42310", "SF42338", "SF42346", "SF42336", 
                 "SF42368", "SF42330", "SF42350", "SF42324")
name_state <- c("CDMX", "EDO_MEX", "NL", "JAL", "VER",  
                "GTO", "PUE", "CHIS")
title_state <- "Number of ATMs by State"
my_start <- '2010-01-01'
my_y <- "Number of ATMs"
my_x <- "Quarterly"

# run the function
df_state <- sie_function(serie_state, name_state, 
            title_state, route="../img/",
            y_lab = my_y, x_lab = my_x,
            startDate=my_start)
```

### 4. Retail payment systems transactions
Information from retail payments at ATMs, Points of Sales (POS), Checks, and Transfers by Electronic Payments, Internet Banking and Phone, from 2016 to 2019.
```{r retail transactions}
serie_trans <- c("SF62275", "SF62278", "SF61610", "SF60841",
                 "SF60842", "SF60843")
name_trans <- c("ATM", "POS", "Check", 
                "Transfer by Electronic Banking", 
                "Transfer by Internet Banking", 
                "Transfer by Phone")
title_trans <- "Retail transactions"
my_y <- "Millions of Pesos"
my_x <- "Quarterly"
my_start <- '2016-01-01'
my_end <- '2019-12-31'

# run the function
df_trans <- sie_function(serie_trans, name_trans, 
            title_trans, route="../img/",
            y_lab = my_y, x_lab = my_x,
            startDate=my_start, endDate=my_end)
```

### 5. Consumer Price Index (INPC)
Main time series of the Consumer Price Index (INPC for it's acronym in spanish), and their core and non-core subindexes. Time window of Enrique Peña Nieto's Presidency of Mexico (December 1, 2012 – November 30, 2018).
```{r inpc}
serie_inpc <- c("SP74625", "SP74626", "SP74628", "SP56337", 
                "SP74631")
name_inpc <- c("INPC (main)", "Merchandise (sub, core)", 
               "Services (sub, core)", 
               "Agricultural (sub, non-core)", 
               "Energy (sub, non-core)")
title_inpc <- "Consumer Price Index"
my_y <- "Index"
my_x <- "Month"
my_start <- '2012-12-01'
my_end <- '2018-11-30'

# run the function
df_inpc <- sie_function(serie_inpc, name_inpc, 
            title_inpc, route="../img/",
            y_lab = my_y, x_lab = my_x,
            startDate=my_start)
```
