---
title: "SIE example with Counterfeit Banknotes and Coins series"
author: "Victor Cuspinera"
date: "22/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the example shared in the __Usage__ section of the main introduction of this Repository.

## Overview

To show the usage of the SIE API with `siebanxicor` R-package, we will go through an example using the time series for the of the __Annual counterfeit domestic banknotes detected for the current denomination in circulation__: 20, 50, 100, 200, 500 and 1000 pesos  (series __SM1249__, __SM1250__, __SM1251__, __SM1252__, __SM1253__, __SM1254__).

[Click here](https://www.banxico.org.mx/SieAPIRest/service/v1/doc/catalogoSeries) to look for the complete catalogue of the SIE's time series published by Banco de Mexico.

### 1. Load library

After the `siebanxicor` package is installed, load this library and also the tidyverse library.
```{r load libraries}
library("siebanxicor")
library("tidyverse")
```


### 2. Use `setToken(token)`

Bring your token and open the SIE API channel with the `setToken` utility function.
```{r bring and set token, warning=FALSE}
# bring the token
token_file <- read.csv("../token/SIE_Token.csv", header=FALSE)

# set the token
setToken(token_file$V2)
```
_Notes:_  
- _If you don't have a token to use SIE API, [**click here**](https://www.banxico.org.mx/SieAPIRest/service/v1/token) to access the official website and obtain one._  
- _I add a csv file where users should paste and save their token to run this example._  


### 3. Get data with `getSeriesData(series, startDate, endDate)`

Get the time series of interest, in this case the series of the annual counterfeit of mexican banknotes per denomination, using the `getSeriesData` function.
```{r getting series}
# setting the variables
my_series <- c("SM1249", "SM1250", "SM1251", "SM1252", "SM1253", "SM1254")
my_start <- '2015-01-01'
my_end <- Sys.Date() #looks for today's date

# getting the series
series <- getSeriesData(my_series, my_start, my_end)
```

...this is the vector returned as result:
```{r getSeriesData results, echo=FALSE}
series
```
_Note: to use the `getSeriesData` function, you should previously call `setToken`._


### 4. Get the metadata with `getSeriesMetadata(series, locale)`

This function returns the general information of series. To select the language of the metadata, set the _locale_ variable as "en" for English, and "es" for Spanish.

```{r getting metadata}
# getting the metadata
getSeriesMetadata(my_series, locale="en")
```

_Note: to use the `getSeriesMetadata` function, you should previously call `setToken`._

### 5. Get a data frame of one series using `getSerieDataFrame(series, idSerie)`
This function will be helpful to get a data frame for the annual counterfeit number of 500 pesos banknotes (__SM1253__ series), from the vector returned by the `getSerieDataFrame` in the previous point #3.

```{r get SM1253 df}
# getting the series
df_SM1253 <- getSerieDataFrame(series, "SM1253")
```

...this is the data frame that we get as result:
```{r results SM1253 df, echo=FALSE}
df_SM1253
```

_Note: to use the `getSerieDataFrame` function, you should previously call `setToken` and `getSerieData`._


### 6. Get the last value of one or more series with `getSeriesCurrentValue(series)`

To get the last value of the annual counterfeit banknotes per denomination series, we will use the `getSeriesCurrentValue` function.

```{r last value}
series_last <- getSeriesCurrentValue(my_series)
```

...this is the data frame taht we get as result after using `getSeriesData`:
```{r getSeriesCurrentValue results}
series_last
```
_Note: to use the `getSeriesCurrentValue` function, you should previously call `setToken`._


### 7. Use the custome function `SIE_function(series_codes, series_names, title_plot, x_lab, y_lab, startDate, endDate, route)`

This function prints the metadata for the Annual Counterfit Banknotes series, prints and saves and their plot, and returns a data frame with these series in tidy format.

```{r custome function}
# call the customed function from an RScript
source("SIE_function.R")

# setting the variables
my_series <- c("SM1249", "SM1250", "SM1251", "SM1252", "SM1253", "SM1254")
my_names <- c("20", "50", "100", "200", "500", "1000")
my_title <- "Annual Counterfeit Banknotes per Denomination"
my_start <- '2010-01-01'
my_end <- Sys.Date() #looks for today's date

# run the function
df <- sie_function(my_series, my_names, my_title, route="../img/",
             y_lab="Number of Pieces", x_lab="Year", startDate=my_start)
```

...this is the data frame that we get as result:
```{r custome function df, echo=FALSE}
df %>% head()
```
